---
title: "Group 1 Project - Ritcher"
format: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 12,
  fig.height = 12
)
```

# Phase 1: Data Wrangling

First stage is to download our various .csv files and perform a join to create one master dataframe. We must match by building_id.

```{r load packages}
#Loading all needed packages
library(tidyverse)
library(DT)
library(patchwork)
library(tidymodels)
```

## Joining the two data sources

```{r join data sources}
# Read in the two data sources
train_labels <- readr::read_csv("train_labels.csv")
train_values <- readr::read_csv("train_values.csv")

#Use Dplyr's left_join to join train_labels and train_values on building_id
raw_data <- left_join(train_values, train_labels, by= "building_id")
```

We now have a raw data set that contains building values AND their associated damage grade.

## Fixing Variable Types

```{r glimpse raw}
glimpse(raw_data)
```

Almost all data is in dbl format, not int. If it's not in double format, it is in character format. These should be factors. Let's start by mutating the columns we need into integers. We will create a new dataframe called "data_cleaned" where all future transformations will occur on it.

```{r mutate to int char}
data_cleaned <- raw_data %>%
  mutate(
    across(
      c(
        count_floors_pre_eq,
        age,
        area_percentage,
        height_percentage,
        damage_grade,
        count_families
      ),
      as.integer
    )
  )

data_cleaned <- data_cleaned %>%
  mutate(
    across(
      c(
        geo_level_1_id, 
        geo_level_2_id,
        geo_level_3_id
      ),
      as.character
    )
  )

```

Looking at the project's data dictionary, we have the rest of the data as binary and/or categorical variables of multiple levels. We will need to make sure both are converted to factors.

```{r mutate factors}
#Parse out data into the last two types, categorical and binary

# Categorical variables (multiple levels)
cat_vars <- c(
  "land_surface_condition",
  "foundation_type",
  "roof_type",
  "ground_floor_type",
  "other_floor_type",
  "position",
  "plan_configuration",
  "damage_grade",
  "legal_ownership_status"
)

# Binary variables (0,1)
binary_vars <- c(
  "has_superstructure_adobe_mud",
  "has_superstructure_mud_mortar_stone",
  "has_superstructure_stone_flag",
  "has_superstructure_cement_mortar_stone",
  "has_superstructure_mud_mortar_brick",
  "has_superstructure_cement_mortar_brick",
  "has_superstructure_timber",
  "has_superstructure_bamboo",
  "has_superstructure_rc_non_engineered",
  "has_superstructure_rc_engineered",
  "has_superstructure_other",
  "has_secondary_use",
  "has_secondary_use_agriculture",
  "has_secondary_use_hotel",
  "has_secondary_use_rental",
  "has_secondary_use_institution",
  "has_secondary_use_school",
  "has_secondary_use_industry",
  "has_secondary_use_health_post",
  "has_secondary_use_gov_office",
  "has_secondary_use_use_police",
  "has_secondary_use_other"
)

# Transform data types
data_cleaned <- data_cleaned %>%
  mutate(
    
    # Convert categorical columns to factors
    across(all_of(cat_vars), as.factor),
    # Convert binary columns to factors (Make sure to keep 0 and 1)
    across(all_of(binary_vars), as.factor))
```

## Finding and Dealing with NAs

Doing a quick skim of the data on the data viewer, there doesn't seem to be any NAs or values that would indicate NAs, but we double check that with code.

```{r check for na}
na_cols<-colSums(is.na(data_cleaned))
na_cols[na_cols>0] %>% sum()
```

We confirmed that there are no NAs and now we check for duplicates.

```{r check for dupes}
# Duplicate rows
data_cleaned %>%
  duplicated() %>%
  sum()
```

We have no NAs and no duplicates.

## Dropping Unnecessary Columns

Building_id will be dropped from the table due to the fact it is just an identifier for each row in the data with no additional value.

```{r drop unnecessary}
# Drop building_id from raw data, as it's just an identifier
data_cleaned <- data_cleaned %>%
  select(-building_id)
```

## Summary of Initial Data Wrangling

We uploaded the two CSV files, performed a left join and saved the new joined data into a dataframe and csv file. We then ensured that the joined data, then called 'raw_data', had all the variables in the correct format. We also confirmed there were no NAs or duplicates. The current main dataframe we will be using is called "**data_cleaned**".

# Phase 2: EDA

Once we have our data into a cleaner place, we can start with performing EDA to get a better feel for the overall data. I think the best way to cut through nearly 40 columns is by splitting our work into different parts.

## `damage_grade` Investigation

We know that `damage_grade` will be the response variable of interest. So let's take a look at the class distribution of the 3 damage types.

```{r plot damage_grade}
# Plot target distribution
ggplot(data_cleaned, aes(damage_grade, fill = damage_grade)) +
  geom_bar() +
  labs(
    title = "Distribution of Damage Grade",
    x = "Damage Grade",
    y = "Count"
  ) +
    scale_fill_manual(
            values = c(
              "1" = "skyblue",
              "2" = "darkgoldenrod1",
              "3" = "red1"
            )
          )
```

So here we see we have some class imbalance here. The damage grade of 2 is the most common damage grade, making just under 150,000 records. Then it was followed by damage level 3 and then damage level 1 trailing far behind. Class imbalance could cause some issues for models in the future.

## Numerical Variable Data Exploration

Let's start by first looking at our purely numerical data. Let's only save the numerical columns to a subset and then generate some summary statistics.

```{r numeric predictors}
#Using our int transformation list as a basis, but we do not need building_id, any of the geo_lists, or building_damage.

int_subset <- c(
  "age",
  "count_floors_pre_eq",
  "area_percentage",
  "height_percentage",
  "count_families"
)

```

Now we can generate the summary stats

```{r review numeric predictors}
data_cleaned %>%
  select(all_of(int_subset)) %>%
  summary()
```

We see some evidence of skew and outliers in this dataset, let's further explore with some histograms.

### Numerical Data Histograms

```{r numeric histograms}
data_cleaned %>%
  select(all_of(int_subset)) %>%
  pivot_longer(everything(),
               names_to = "variable",
               values_to = "value") %>%
  ggplot(aes(value)) +
  geom_histogram(bins = 40, 
                 fill = "purple",
                 color = "white") +
  facet_wrap(~variable, scales = "free") +
  theme_minimal()

```

### Numerical Data Boxplots

To further investigate our numerical data, a boxplot is applied to three of the variables, `age` `area_percentage`, and `height_percentage`. The count variables were excluded from this analysis due to the variables having relatively few distinct values, making the boxplots illegible.

```{r numeric boxplots}

#Noted some wild outliers so let's look at boxplot. Not bothering with select(all_of)

age_boxplot <- ggplot(data_cleaned, aes(x = age)) +
  geom_boxplot(fill = "purple", color = "black") +
  theme_minimal() +
  labs(
    title = "Boxplot of Age",
    x = "Age"
  )

area_boxplot <- ggplot(data_cleaned, aes(x = area_percentage)) +
  geom_boxplot(fill = "purple", color = "black") +
  theme_minimal() +
  labs(
    title = "Boxplot of Area Percentage",
    x = "Area Percentage"
  )

height_boxplot <- ggplot(data_cleaned, aes(x = height_percentage)) +
  geom_boxplot(fill = "purple", color = "black") +
  theme_minimal() +
  labs(
    title = "Boxplot of Height Percentage",
    x = "Height Percentage"
  )

  (age_boxplot / area_boxplot / height_boxplot) +
  plot_annotation(
    title = "Boxplots of Some Numeric Variables"
  )
```

**Brief summary of findings:**

`age` - Age skews very young in the dataset. The data is extremely right-skewed, meaning that most buildings in Nepal are only a few years old. There are very few buildings that are older. It can be hard to see in this graph, but there are a few buildings that are "995" years old. Is that correct? Usually an entry code of 999 would indicate an NA value, but 995 could be a realistic age of building in Nepal. After some brief googling, there are some buildings in Nepal that come from 17 CE, or even older. https://www.blackgate.com/2015/04/29/remembering-nepals-ancient-cultural-heritage/. Average age is 26 years, while the median is 16. My suggestion is to remove the extreme outliers from the data and log transform the rest.

`area_percentage` - This is the normalized area of the building's footprint. Appropriately, this data seems to be already transformed to be roughly normal to very moderately right-skewed. Most values are less than 100 units (probably meters). Mean is 8.018 with median being slightly lower at 7. I recommend log transform.

`count_families` - This is a distinct variables that counts the amount of families that lived in the building. Most of these values hover at around 1 to 2 families in a building, with some buildings being unoccupied by families (note zero values). There are some buildings that hold more families, but these are likely outliers. They don't even show up on the histogram. Mean and median are roughly equal with a mean of 0.9839 and a median of 1.

`count_floors_pre_eq` - This variable is the floor count in the building before the building. We can see that the majority of structures pre-earthquake were two stories. This has a rough normal curve, but structures greater than 3 stories are relatively rare. Mean and median are roughly the same with a median of 2.00 and a mean of 2.13.

`height_percentage` - Here we see the normalized height of a building's footprint. We can clearly see that the normalization worked, as the histogram is roughly in a normal with a moderate skew. The mean is roughly 5.424 while the median is 5. I recommend log transform.

We also see that all of these variables have some level of outliers. The worst case is age, as mentioned before. We have some structures that appear to be almost a millennium old! For the most extreme outliers, I would suggest removing them.

### Numerical Data Correlation Matrix

We see how each numerical variable interacts with one another in isolation, but how do they act together?

```{r numeric corr matrix}
corr_df <- data_cleaned %>%
  select(all_of(int_subset))

#Make a heatmap!
GGally::ggcorr(
  corr_df,
  label = TRUE,
  label_round = 2
)
```

As we can clearly see, we have relatively LOW correlation between each numerical variable except one pair. `count_floors_pre_eq` and `height_percentage` are highly positively correlated. When we start our modeling, we may need to drop one of these columns for another.

### Numerical Values and `damage_grade`

```{r numerics against damage_grade}
data_cleaned %>%
  pivot_longer(
    cols = all_of(int_subset),
    names_to = "variable",
    values_to = "value"
  ) %>%
  ggplot(aes(damage_grade, value, fill = damage_grade)) +
  geom_boxplot() +
  scale_fill_manual(
            values = c(
              "1" = "skyblue",
              "2" = "darkgoldenrod1",
              "3" = "red1"
            )
          ) +
  facet_wrap(~variable, scales = "free") +
  theme(legend.position = "none") +
  labs(
    title = "Numeric Variables vs Damage Grade"
  )
```

Okay so when we parse out the numerical data across damage grades, we see that `area_percentage`, `count_families`, and `age` so not differ radically from one another. However, we do see differences in `count_floors_pre_eq` and `height_percentage`. Knowing that these two were highly correlated with one another, this makes sense. It appears that the overall height of the building matters when predicting its level of destruction in the earthquake. This could be an important predictor!

## Categorical Variable Data Exploration

We now move onto looking at the categorical and binary variables found within this dataset. I want to not only look at their distribution, but also see how it parses out with the damage grades.

```{r categorical review}
#proportion by damage_grade
cat_plot_grid <-
  wrap_plots(
    # go through every variable (var) found in cat_vars (doesn't include binary vars)
    lapply(cat_vars, function(var) {
      ggplot(
        data_cleaned,
        aes_string(x = var, fill = "damage_grade") #Remember fill is needed for proportion bar charts
      ) +
        geom_bar(
          position = "fill", 
          color = "white",
          linewidth = 0.2
        ) +
          scale_fill_manual(
            values = c(
              "1" = "skyblue",
              "2" = "darkgoldenrod1",
              "3" = "red1"
            )
          ) +

        scale_y_continuous(
          labels = scales::percent_format()
        ) +

        labs(
          title = var,
          x = NULL,
          y = "Proportion"
        ) +

        theme_minimal(base_size = 10) + #No chartjunk - Tufte's book. 

        theme(
          plot.title = element_text(
            face = "bold",
            size = 10,
            hjust = 0.5
          ),
          # Fix up the x-axis labels within each grid cell
          axis.text.x = element_text(
            angle = 45,
            hjust = 1,
            size = 10
          ),
          # Fix up our y axis labels and our legend
          axis.text.y = element_text(size = 7),

          legend.position = "bottom",

          legend.title = element_blank(),

          panel.grid.minor = element_blank()
        )

    }),

    ncol = 3,
    guides = "collect"

  ) +

  plot_annotation(
    title = "Categorical Variables vs Damage Grade",
    theme = theme(
      plot.title = element_text(
        size = 16,
        face = "bold",
        hjust = 0.5
      )
    )
  )

# Display patchwork grid
cat_plot_grid
```

**Finding per above variables:**

`land_surface_condition` - While damage grade 2 seems to dominate all three surface conditions, it appears that the proportions are relatively similar for each condition type. Condition type 'o' seems to have a higher proportion of major damage, while condition 't' shares slightly higher damage level 1. However, the patterns look shockingly similar.

`foundation_type` - Foundation type really seems to matter here. It appears that foundation type 'l' seems to disproportionally have little damage levels (damage level 1) compared to the other foundation types. Meanwhile, foundation type 'r' seems to show the most damage, having a high proportion of 3 values and low proportion of 1 values. This looks like an important variable.

`roof_type` - Roof type also shows quite a bit of differences. For example, roof type 'x' seems to have quite a bit of resilience, with a lot of 1 and 2 values for damage. Meanwhile, the other two roof types show similar proportions.

`ground_floor_type` - We see another bit of variation among ground floor types. Ground floor type 'v' shows less destruction compared to the others, while floor type 'f' and "x" see a lot more. This is probably an important variable.

`other_floor_type` - We see more variation between other floor types, with floor type 's' having a higher proportion of low damage compared to the other floor types. Meanwhile, floor type 'q' seems to have a higher proportion of high to medium damage. Another important variable.

`position` - Position seems to have somewhat similar proportions of damage between position type. Perhaps a moderately important variable.

`plan_configuration` - Plan configuration sees some variation. It's not as extreme as foundation type or roof type, but variation is most noticeable with the low damage grades and highest damage grades.

`legal_ownership_status` - Legal ownership status DOES see some differences among the groups. Namely that ownership type 'a' saw the highest proportion of level 1 damage while ownership type 'w' saw the least. This could be another important variable.

## Binary Variable Data Exploration

This section compares the binary variables for superstructure and secondary use.

```{r super and secondary}
# Make damage_grade a factor
data_cleaned <- data_cleaned %>%
  mutate(damage_grade = factor(damage_grade))

# Get the binary variable groups
superstructure_vars <- binary_vars[str_detect(binary_vars, "has_superstructure")]
secondary_use_vars <- binary_vars[str_detect(binary_vars, "has_secondary_use")]
```

```{r review superstructure}
#Review each superstructure variable by itself - helps show which building materials had much or less severe damage

superstructure_summary <- data_cleaned %>%
  select(damage_grade, all_of(superstructure_vars)) %>%
  pivot_longer(
    cols = all_of(superstructure_vars),
    names_to = "variable",
    values_to = "present"
  ) %>%
  filter(present == 1) %>%
  count(variable, damage_grade) %>%
  group_by(variable) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

ggplot(superstructure_summary,
       aes(x = variable, y = prop, fill = damage_grade)) +
  geom_col(color = "white") +
  scale_fill_manual(
    values = c(
      "1" = "skyblue",
      "2" = "darkgoldenrod1",
      "3" = "red1"
    )
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
labs(
  title = "Damage Levels for Different Building Materials",
  x = "Building Material",
  y = "Percent of Buildings",
  fill = "Damage Grade"
) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Here we can clearly see that each building material contains different proportions for building damage type. Some, such as stone buildings, contain a much higher proportion of damage level 3 damage than others. RC_Engineered buildings seem to have the least amount of damage. However, trying to keep track of every binary building material type is difficult, causing some issues for interpretation later. In addition, it is hard to distinguish between variables that have similar, yet still different, structures.

```{r plot secondary use}
# Plot detailed secondary use variables only - exclude has_secondary_use because it is the broad master flag

secondary_summary <- data_cleaned %>%
  select(damage_grade, all_of(setdiff(secondary_use_vars, "has_secondary_use"))) %>%
  pivot_longer(
    cols = -damage_grade,
    names_to = "variable",
    values_to = "present"
  ) %>%
  filter(present == 1) %>%
  count(variable, damage_grade) %>%
  group_by(variable) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

ggplot(secondary_summary,
       aes(x = variable, y = prop, fill = damage_grade)) +
  geom_col(color = "white") +
  scale_fill_manual(
    values = c(
      "1" = "skyblue",
      "2" = "darkgoldenrod1",
      "3" = "red1"
    )
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
labs(
  title = "Damage Grade by Secondary Use",
  x = "Secondary Use",
  y = "Percent of Buildings",
  fill = "Damage Grade"
) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Converting to a Binary Classification Problem

After performing EDA and considering different model approaches, we've decided to restructure this into a binary classification problem which classifies whether or not there would be damage. We've grouped levels 2 and 3 of damage_grade as this simplifies our classification problem in an appropriate manner given the ordinal nature of our classes. We determined it would be most important to predict simply whether there would be damage, instead of whether that damage was moderate to high, as understanding buildings which are at risk for damage is the key.

- **damaged:** 0 for damage_grade=1 (minimal)

- **damaged**: 1 for damage_grade=2,3 (moderate to high)

Below we see the new predictor and its distribution.

```{r add new damaged indicator and plot distribution}
data_cleaned <- data_cleaned |>
  mutate(damaged = factor(if_else(damage_grade %in% c(2, 3), "Yes", "No"), 
                          levels = c("Yes", "No")))

#remove old damage grade column
data_cleaned <- data_cleaned %>%
  select(-damage_grade)
```

```{r plot new damaged indicator distribution}
# Plot target distribution
ggplot(data_cleaned, aes(damaged, fill = damaged)) +
  geom_bar() +
  labs(
    title = "Distribution of Damaged",
    x = "Damaged",
    y = "Count"
  ) +
    scale_fill_manual(
            values = c(
              "No" = "skyblue",
              "Yes" = "red1"
            )
          )
```

# Test and Train Split

Our next steps are to split the data into an 80/20 test-train split using *damaged* for strata to ensure that each data set (test and train) have equal proportions of our damaged indicator.

```{r stratify split on damaged}
set.seed(420) 
data_split <- initial_split(data_cleaned, prop = 0.80, strata=damaged) 
train <- training(data_split) 
test <- testing(data_split)

#Setting up our resamples for cross-val using training set
resamples <- vfold_cv(train, v = 5, strata = damaged)
```

```{r lasso recipe prep}
# groups geographic levels together before dummy encoding

lasso_recipe <-
  recipe(damaged ~ ., data = train) %>%
#group rare categories <1% into "other"
#to reduce dummy variables/high cardinality
  step_other(
    geo_level_2_id,
    geo_level_3_id,
    threshold = 0.01
  ) %>%
#convert categorical to dummy numerics
  step_dummy(all_nominal_predictors()) %>%
#removes any variables without variance
#which indicates no model value
  step_zv(all_predictors()) %>%
#scaled numerics predictors
  step_normalize(all_numeric_predictors())
```

```{r lasso predictor check}
# Check # of predictors after preprocessing
# To verify that dummy encoding & grouping of rare
# geographic levels did not create any large feature space

prep(lasso_recipe) %>%
  juice() %>%
  dim()
```

```{r lasso penalty grid}
# Create a grid of penalty values to check
# The penalty controls how strongly LASSO shrinks coefficients toward 0
# Larger penalties result in larger feature selection

lasso_grid <-
  grid_regular(
    penalty(),
    levels = 10
  )
```

```{r lasso specs}
# LASSO logistic reg model
# Use mixture = 1 for L1 regularization, which can shrink
# coefficients to 0 and do automatic feature selection

library(glmnet)

lasso_spec <-
  logistic_reg(
    penalty = tune(), #seek best value by tuning
    mixture = 1
  ) %>%
  set_engine("glmnet")
```

```{r lasso wf}
# Combine the preprocessing recipe/ model spec into a workflow.

lasso_wf <-
  workflow() %>%
  add_recipe(lasso_recipe) %>%
  add_model(lasso_spec)
```

```{r parallel computing}
# Make faster
library(doParallel)

cl <- makePSOCKcluster(parallel::detectCores() - 1)
registerDoParallel(cl)
```

```{r tune lasso penalty}
# Tune the penalty parameter using X val
# Evaluate model perf across penalty values
# and select the value that max ROC AUC.

set.seed(420)

lasso_tuned <-
  tune_grid(
    lasso_wf,
    resamples = resamples, #5 CFV
    grid = lasso_grid, #10 penalty levels
    metrics = metric_set(roc_auc, bal_accuracy)
  )
```

```{r plot tuning of lasso}
# Visualize how ROC AUC changes
# as the penalty parameter get bigger

autoplot(lasso_tuned)
```

```{r get best penalty}
# Select the penalty value that produced the highest ROC AUC

best_penalty <- select_best(
  lasso_tuned,
  metric = "roc_auc"
)

best_penalty
```

```{r fit lasso with best penalty}
# Fit the LASSO model using the best penalty value

final_lasso_wf <-
  finalize_workflow(
    lasso_wf,
    best_penalty
  )

final_lasso_fit <-
  fit(
    final_lasso_wf,
    data = train
  )
```

```{r}
# Get coefficient estimates from the model at best penalty value
coef_table <-
  tidy(
    extract_fit_parsnip(final_lasso_fit)
  )
```

```{r show predictors by coeff mag}
# Display the predictors ordered by coefficient magnitude

coef_table %>%
  filter(term != "(Intercept)") %>%
  arrange(desc(abs(estimate)))%>%
  filter(estimate != 0)
```

```{r get count of predictors with nonzero coeff}
# Count the # of coefficients in the model
# which have not been removed/shrunk to 0
coef_table %>%
  filter(estimate != 0)%>%
  nrow()
```

```{r show which predictors are dropped}
# show which have  been removed/shrunk to 0
coef_table %>%
filter(estimate == 0)
```


## LASSO Feature Selection Results
 ROC AUC: 0.895
 Best penalty: 1e-10
 Predictors retained: 90

Conclusion: LASSO provided minimal feature reduction; most predictors appear to contain useful predictive information. Even when selecting a higher penalty value, such as 10^-3, it resulted in minimal feature selection. We suspect this is because our data violates some of the assumptions of LASSO logistic regression; namely our sample size is much larger than our parameters, and many of our features are nested within one another and are not mutually exclusive, because of that, we will attempt to use feature importance from random forest models to engage in feature selection. 

## Attempt 2: Feature Importance Measures from RF Models

Based on the following tutorial: https://gedeck.github.io/machine-learning-with-tidymodels/src/605-feature-importance.html

Note that I chose permutation as our importance metric as it is model agnostic.

```{r}
library(tidymodels)
library(ranger)
library(vip)
library(important)

#Double check this is actually as factors and not dummies

set.seed(420)

data_rf <- train %>%
  mutate(
    damaged = as.factor(damaged),
    across(
      c(
        geo_level_1_id,
        geo_level_2_id,
        geo_level_3_id,
        land_surface_condition,
        foundation_type,
        roof_type,
        ground_floor_type,
        other_floor_type,
        position,
        plan_configuration,
        legal_ownership_status
      ),
      as.factor
    )
  )

#recipe for random forest model
rf_recipe <-
  recipe(damaged ~ ., data = data_rf)

#pretty forest
rf_spec <-
  rand_forest(
    trees = 100
  ) %>%
  set_engine(
    "ranger",
    importance = "permutation"
  ) %>%
  set_mode("classification")

# workflow as usual
rf_wf <-
  workflow() %>%
  add_recipe(rf_recipe) %>%
  add_model(rf_spec)

# Fit the model

rf_fit <-
  rf_wf %>%
  fit(data = data_rf)

# Generate variable importance plot to determine what to keep
rf_fit %>%
  extract_fit_engine() %>%
  vip(num_features = 20)

```

Here we see that Geo level 1 dominates all other variables, followed by mud_mortar_stone, and other geo_levels. I suspect the nested nature of geo levels is dominating the feature importance random forest model, so let's remove the finer granularity of geo_level_2 and geo_level_3 and keep ONLY geo_level_1.Let's also remove count_floors_pre_eq as it is highly correlated with height_percentage. Let's rerun my random forest and keep ONLY geo_level_1 in that training:

```{r}
#code from various documents/tutorials
set.seed(420)
# Recipe with useless variables removed
rf_recipe <-
  recipe(damaged ~ ., data = train) %>%
  step_rm(
    geo_level_2_id,
    geo_level_3_id,
    count_floors_pre_eq
  ) %>%
  step_zv(all_predictors()) #from one of the tutorials 

# Random Forest with permutation importance
rf_spec <-
  rand_forest(
    trees = 200
  ) %>%
  set_engine(
    "ranger",
    importance = "permutation",
    num.threads = parallel::detectCores() - 1 #speeds up
  ) %>%
  set_mode("classification")

# Workflow
rf_wf <-
  workflow() %>%
  add_recipe(rf_recipe) %>%
  add_model(rf_spec)

# Fit ONE model
final_rf_fit <-
  rf_wf %>%
  fit(data = train)

# Variable importance plot
final_rf_fit %>%
  extract_fit_engine() %>%
  vip(num_features = 20)

# numeric importance table for all variables
importance_tbl <-
  final_rf_fit %>%
  extract_fit_engine() %>%
  vip::vi()

importance_tbl %>%
  arrange(desc(Importance))

```


```{r}

importance_tbl <- importance_tbl %>%
  arrange(desc(Importance)) %>%
  mutate(
    cum_importance = cumsum(Importance) / sum(Importance)
  )

selected_features <- importance_tbl %>%
  filter(cum_importance <= 0.90) %>%
  pull(Variable)

selected_features <- c(
  selected_features,
  importance_tbl %>%
    filter(cum_importance > 0.90) %>%
    slice(1) %>%
    pull(Variable)
)
selected_features <- unique(selected_features)

train_reduced <- train %>%
  select(any_of(selected_features), damaged)

test_reduced <- test %>%
  select(any_of(selected_features), damaged)

```

So, in summary, we used a Random Forest Classification Model (using training data) and calculated permutation based variable importance scores. Prior to this, we removed hierarchical variables `geo_level_2_id` and `geo_level_3_id`. This was because geo_level_1_id was one of the most important measures for location and was the lowest granularity geographic variable, making geo levels 2 and 3 largely redundant. We also removed count_floors_pre_eq because it was strongly correlated with height_percentage. Finally, we selected variables that were most important and accounted for 90% of total importance in the data. This preserves most of the future model's predictive performance while simplifying model interpretation and workflow.


# Helper Functions

We establish the set of metrics to evaluate each model against to maintain consistency and use helper functions to ensure accurate calculation against each model with minimal code redundancy.

```{r define metric set and helper functions}
#define metric set to be used across models for evaluations and comparison
class_metrics<- metric_set(roc_auc, bal_accuracy, sens, spec, f_meas, kap)

#pass fitted model using best parameters + model name
#function will predict against test data
get_model_preds <- function(fitted_model, model_name) {
  augment(fitted_model, new_data = test)
}

#separate visuals out for organization
get_model_plots <- function(aug_data, model_name) {
  roc_plot<-aug_data |>
    roc_curve(truth = damaged, .pred_Yes) |>
    autoplot() +
    labs(
      title = paste0(model_name, " ROC Curve"),
      x = "False Positive Rate",
      y = "True Positive Rate"
    )
}

#usage example: 
#get_model_metrics(final_lasso_fit,"Lasso")
get_model_metrics <- function(fitted_model, model_name) {

  #get predicted model data to send to graphs
  aug_data<-get_model_preds(fitted_model, model_name)

  #use same table for metrics table
  metrics_results<-aug_data|>
    class_metrics(truth = damaged, estimate = .pred_class, .pred_Yes) |>
    mutate(model_name = model_name) |>
    select(model_name, .metric, .estimate) |>
    pivot_wider(names_from = .metric, values_from = .estimate)
   
   #show graphics
   print(get_model_plots(aug_data, model_name)) 
   return(metrics_results)
}
```

# LASSO Evaluation

```{r lasso evaluation}
get_model_metrics(final_lasso_fit,"Lasso")|>
  knitr::kable(digits = 4)
```
