---
title: "Group 1 Project - Ritcher"
format: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 12,
  fig.height = 12
)
```

# Phase 1: Data Wrangling

First stage is to download our various .csv files and perform a join to create one master dataframe. We must match by `building_id`.

```{r load packages}
#Loading all needed packages
#include parallel computing
library(tidyverse)
library(DT)
library(patchwork)
library(tidymodels)
library(future)
plan(multisession, workers = parallel::detectCores(logical = FALSE))
```

## Joining the two data sources

Below we create a data set that contains building values AND their associated damage grade by joining the files based on their `building_id`.

```{r join data sources}
# Read in the two data sources
train_labels <- readr::read_csv("train_labels.csv")
train_values <- readr::read_csv("train_values.csv")

#Use Dplyr's left_join to join train_labels and train_values on building_id
raw_data <- left_join(train_values, train_labels, by= "building_id")
```

## Fixing Variable Types

```{r glimpse raw}
glimpse(raw_data)
```

[We begin by addressing a few data type issues:]{.underline}

- double values that should be integers

- character values that should be factors

- binary and categorical variables of multiple levels that should be factors

We will create a new dataframe called **`data_cleaned`** where all future transformations on our raw data will occur.

```{r mutate to int char}
data_cleaned <- raw_data %>%
  mutate(
    across(
      c(
        count_floors_pre_eq,
        age,
        area_percentage,
        height_percentage,
        damage_grade,
        count_families
      ),
      as.integer
    )
  )

data_cleaned <- data_cleaned %>%
  mutate(
    across(
      c(
        geo_level_1_id, 
        geo_level_2_id,
        geo_level_3_id
      ),
      as.character
    )
  )

```

```{r mutate cat and binary into factors}
# Categorical variables (multiple levels)
cat_vars <- c(
  "land_surface_condition",
  "foundation_type",
  "roof_type",
  "ground_floor_type",
  "other_floor_type",
  "position",
  "plan_configuration",
  "damage_grade",
  "legal_ownership_status"
)

# Binary variables (0,1)
binary_vars <- c(
  "has_superstructure_adobe_mud",
  "has_superstructure_mud_mortar_stone",
  "has_superstructure_stone_flag",
  "has_superstructure_cement_mortar_stone",
  "has_superstructure_mud_mortar_brick",
  "has_superstructure_cement_mortar_brick",
  "has_superstructure_timber",
  "has_superstructure_bamboo",
  "has_superstructure_rc_non_engineered",
  "has_superstructure_rc_engineered",
  "has_superstructure_other",
  "has_secondary_use",
  "has_secondary_use_agriculture",
  "has_secondary_use_hotel",
  "has_secondary_use_rental",
  "has_secondary_use_institution",
  "has_secondary_use_school",
  "has_secondary_use_industry",
  "has_secondary_use_health_post",
  "has_secondary_use_gov_office",
  "has_secondary_use_use_police",
  "has_secondary_use_other"
)

# Transform data types
data_cleaned <- data_cleaned %>%
  mutate(
    # Convert categorical columns to factors
    across(all_of(cat_vars), as.factor),
    # Convert binary columns to factors (Make sure to keep 0 and 1)
    across(all_of(binary_vars), as.factor))
```

## Finding and Dealing with NAs

There does not appear to be any N/As or values that would indicate N/As, but we double check that with code below

```{r check for na}
na_cols<-colSums(is.na(data_cleaned))
na_cols[na_cols>0] %>% sum()
```

We confirmed that there are no NAs and now we check for duplicates.

```{r check for dupes}
# Duplicate rows
data_cleaned %>%
  duplicated() %>%
  sum()
```

We have no NAs and no duplicates so no action is required here.

## Dropping Unnecessary Columns

`building_id` will be dropped from the table due to the fact it is just an identifier for each row in the data with no additional value.

```{r drop unnecessary columns}
# Drop building_id from raw data, as it's just an identifier
data_cleaned <- data_cleaned %>%
  select(-building_id)
```

## Summary of Initial Data Wrangling

We uploaded the two CSV files, performed a left join on `building_id` and saved the new joined data into a dataframe. We then ensured that the joined data, then called **`raw_data`**, had all the variables in the correct format. We also confirmed there were no NAs or duplicates. The current main dataframe with necessary transformations applied to be used for subsequent steps is called **`data_cleaned`**.

# Phase 2: Exploratory Data Analysis (EDA)

We now continue with Exploratory Data Analysis now that we have addressed any obvious issues.

## `damage_grade` Investigation

We know that `damage_grade` will be the response variable of interest, so we review the class distribution of the 3 damage grades.

```{r plot damage_grade}
# Plot target distribution
ggplot(data_cleaned, aes(damage_grade, fill = damage_grade)) +
  geom_bar() +
  labs(
    title = "Distribution of Damage Grade",
    x = "Damage Grade",
    y = "Count"
  ) +
    scale_fill_manual(
            values = c(
              "1" = "skyblue",
              "2" = "darkgoldenrod1",
              "3" = "red1"
            )
          )
```

This exposes a clear class imbalance: `damage_grade = 2` is the most common damage grade, accounting for just under 150,000 records, followed by `damage_grade=3` and then `damage_grade=1` trailing far behind. We must make special considerations to ensure the class imbalance is handled in our modeling.

## Numerical Variable Data Exploration

Let's start by first looking at our purely numerical data. Let's only save the numerical columns to a subset and then generate some summary statistics.

```{r numeric predictors}
#Using our int transformation list as a basis
#but we do not need building_id, any of the geo_lists, or building_damage.

int_subset <- c(
  "age",
  "count_floors_pre_eq",
  "area_percentage",
  "height_percentage",
  "count_families"
)

```

Per the below summary stats, we see some evidence of skew and outliers in this dataset.

```{r review numeric predictors}
data_cleaned %>%
  select(all_of(int_subset)) %>%
  summary()
```

### Numerical Data Histograms

We explore our observations further using histograms.

```{r numeric histograms}
data_cleaned %>%
  select(all_of(int_subset)) %>%
  pivot_longer(everything(),
               names_to = "variable",
               values_to = "value") %>%
  ggplot(aes(value)) +
  geom_histogram(bins = 40, 
                 fill = "purple",
                 color = "white") +
  facet_wrap(~variable, scales = "free") +
  theme_minimal()

```

### Numerical Data Boxplots

To further investigate our numerical data, a boxplot is applied to three of the variables, `age` `area_percentage`, and `height_percentage`. The count variables were excluded from this analysis due to the variables having relatively few distinct values, making the boxplots illegible.

```{r numeric boxplots}

#Noted some wild outliers so let's look at boxplot. Not bothering with select(all_of)

age_boxplot <- ggplot(data_cleaned, aes(x = age)) +
  geom_boxplot(fill = "purple", color = "black") +
  theme_minimal() +
  labs(
    title = "Boxplot of Age",
    x = "Age"
  )

area_boxplot <- ggplot(data_cleaned, aes(x = area_percentage)) +
  geom_boxplot(fill = "purple", color = "black") +
  theme_minimal() +
  labs(
    title = "Boxplot of Area Percentage",
    x = "Area Percentage"
  )

height_boxplot <- ggplot(data_cleaned, aes(x = height_percentage)) +
  geom_boxplot(fill = "purple", color = "black") +
  theme_minimal() +
  labs(
    title = "Boxplot of Height Percentage",
    x = "Height Percentage"
  )

  (age_boxplot / area_boxplot / height_boxplot) +
  plot_annotation(
    title = "Boxplots of Some Numeric Variables"
  )
```

**Summary of numeric variable findings and recommendations based on boxplots:**

- `age` - Age skews very young in the dataset. The data is extremely right-skewed, meaning that most buildings in Nepal are only a few years old. There are very few buildings that are older. It can be hard to see in this graph, but there are a few buildings that are "995" years old. Per <https://www.blackgate.com/2015/04/29/remembering-nepals-ancient-cultural-heritage/>, there are indeed some buildings in Nepal that come from 17 CE, or even older which rules out our theory this may be a data issue. Given the average age is 26 years and the median is 16.

  - [Recommendation:]{.underline} Remove the extreme outliers from the data and log transform the rest. (see below)

  - To avoid issues with skewing the model due to the extreme age outliers, we will remove the rows that have age=995. 1,390 rows are removed, which is less than 1% of the data. See below.

```{r}
#remove the rows that have age=995 as outliers
data_cleaned <- data_cleaned %>% 
  filter(age != 995)
```

- `area_percentage` - This is the normalized area of the building's footprint. This data seems to be already transformed to be roughly normal to very moderately right-skewed. Most values are less than 100 units (probably meters). Mean is 8.018 with median being slightly lower at

  - [Recommendation:]{.underline} Log transform.

- `count_families` - This is a distinct variables that counts the amount of families that lived in the building. Most of these values hover at around 1 to 2 families in a building, with some buildings being unoccupied by families. There are some buildings that hold more families, but these are likely outliers, as they don't even show up on the histogram. Mean and median are roughly equal with a mean of 0.9839 and a median of 1.

  - [Recommendation:]{.underline} no action

- `count_floors_pre_eq` - This variable is the floor count in the building before the damage. We can see that the majority of structures pre-earthquake were two stories. This has a rough normal curve, but structures greater than 3 stories are relatively rare. Mean and median are roughly the same with a median of 2.00 and a mean of 2.13.

  - [Recommendation:]{.underline} no action

- `height_percentage` - Here we see the normalized height of a building's footprint. We can clearly see that the normalization worked, as the histogram is roughly in a normal with a moderate skew. The mean is roughly 5.424 while the median is 5.

  - [Recommendation:]{.underline} Log Transform

## Numerical Data Correlation Matrix

So far we have reviewed numeric predictors in isolation, but how do they behave together?

```{r numeric corr matrix}
corr_df <- data_cleaned %>%
  select(all_of(int_subset))

#Make a heatmap!
GGally::ggcorr(
  corr_df,
  label = TRUE,
  label_round = 2
)
```

As we can clearly see, we have relatively low correlation between each numerical variable except one pair: `count_floors_pre_eq` and `height_percentage` which are highly positively correlated.

- [Recommendation:]{.underline} Drop `count_floors_pre_eq`
- To avoid issues with skewing the model due to highly correlated values, we drop the column below

```{r}
#remove the count_floors_pre_eq column keeping height_percentage because they are highly correlated
data_cleaned <- select(data_cleaned, -count_floors_pre_eq)
int_subset<-int_subset[-2]#removed count_floors_pre_eq since predictor has been dropped
```

### Numerical Values and `damage_grade`

When we parse out the numerical predictors across damage grades, we see that `area_percentage`, `count_families`, and `age` so not differ radically from one another.

However, we do see differences in `height_percentage`. It appears that the overall height of the building matters when predicting its level of destruction in the earthquake. This could be an important predictor!

```{r numerics against damage_grade}
data_cleaned %>%
  pivot_longer(
    cols = int_subset, 
    names_to = "variable",
    values_to = "value"
  ) %>%
  ggplot(aes(damage_grade, value, fill = damage_grade)) +
  geom_boxplot() +
  scale_fill_manual(
            values = c(
              "1" = "skyblue",
              "2" = "darkgoldenrod1",
              "3" = "red1"
            )
          ) +
  facet_wrap(~variable, scales = "free") +
  theme(legend.position = "none") +
  labs(
    title = "Numeric Variables vs Damage Grade"
  )
```

## Categorical Variable Data Exploration

We now move onto exploration of the categorical and binary variables found within this dataset. Our goal is to review their overall distribution as well as their distribution specific to damage grade.

```{r categorical review}
#proportion by damage_grade
cat_plot_grid <-
  wrap_plots(
    # go through every variable (var) found in cat_vars (doesn't include binary vars)
    lapply(cat_vars, function(var) {
      ggplot(
        data_cleaned,
        aes_string(x = var, fill = "damage_grade") #Remember fill is needed for proportion bar charts
      ) +
        geom_bar(
          position = "fill", 
          color = "white",
          linewidth = 0.2
        ) +
          scale_fill_manual(
            values = c(
              "1" = "skyblue",
              "2" = "darkgoldenrod1",
              "3" = "red1"
            )
          ) +

        scale_y_continuous(
          labels = scales::percent_format()
        ) +

        labs(
          title = var,
          x = NULL,
          y = "Proportion"
        ) +

        theme_minimal(base_size = 10) + #No chartjunk - Tufte's book. 

        theme(
          plot.title = element_text(
            face = "bold",
            size = 10,
            hjust = 0.5
          ),
          # Fix up the x-axis labels within each grid cell
          axis.text.x = element_text(
            angle = 45,
            hjust = 1,
            size = 10
          ),
          # Fix up our y axis labels and our legend
          axis.text.y = element_text(size = 7),

          legend.position = "bottom",

          legend.title = element_blank(),

          panel.grid.minor = element_blank()
        )

    }),

    ncol = 3,
    guides = "collect"

  ) +

  plot_annotation(
    title = "Categorical Variables vs Damage Grade",
    theme = theme(
      plot.title = element_text(
        size = 16,
        face = "bold",
        hjust = 0.5
      )
    )
  )

# Display patchwork grid
cat_plot_grid
```

**Summary of categorical variable findings:**

- `land_surface_condition` - While damage grade 2 seems to dominate all three surface conditions, it appears that the proportions are relatively similar for each condition type. Condition type 'o' seems to have a higher proportion of major damage, while condition 't' shares slightly higher damage level 1. However, the patterns look shockingly similar.

- `foundation_type` - Foundation type really seems to matter here. It appears that foundation type 'l' seems to disproportionately have little damage levels (damage level 1) compared to the other foundation types. Meanwhile, foundation type 'r' seems to show the most damage, having a high proportion of 3 values and low proportion of 1 values. This looks like an important variable.

- `roof_type` - Roof type also shows quite a bit of differences. For example, roof type 'x' seems to have quite a bit of resilience, with a lot of 1 and 2 values for damage. Meanwhile, the other two roof types show similar proportions.

- `ground_floor_type` - We see another bit of variation among ground floor types. Ground floor type 'v' shows less destruction compared to the others, while floor type 'f' and "x" see a lot more. This is probably an important variable.

- `other_floor_type` - We see more variation between other floor types, with floor type 's' having a higher proportion of low damage compared to the other floor types. Meanwhile, floor type 'q' seems to have a higher proportion of high to medium damage. Another important variable.

- `position` - Position seems to have somewhat similar proportions of damage between position type. Perhaps a moderately important variable.

- `plan_configuration` - Plan configuration sees some variation. It's not as extreme as foundation type or roof type, but variation is most noticeable with the low damage grades and highest damage grades.

- `legal_ownership_status` - Legal ownership status DOES see some differences among the groups. Namely that ownership type 'a' saw the highest proportion of level 1 damage while ownership type 'w' saw the least. This could be another important variable.

- [Recommendation:]{.underline} No action to be taken on categorical variables yet.

## Binary Variable Data Exploration

This section compares the binary variables for superstructure and secondary use.

```{r super and secondary}
# Make damage_grade a factor
data_cleaned <- data_cleaned %>%
  mutate(damage_grade = factor(damage_grade))

# Get the binary variable groups
superstructure_vars <- binary_vars[str_detect(binary_vars, "has_superstructure")]
secondary_use_vars <- binary_vars[str_detect(binary_vars, "has_secondary_use")]
```

Below we can see that each building material contains different proportions for building damage type. Some, such as stone buildings, contain a much higher proportion of `damage_grade=3` than others. However, trying to keep track of every binary building material type is difficult, causing some issues for interpretation later. In addition, it is hard to distinguish between variables that have similar, yet still different, structures.

```{r review superstructure}
#Review each superstructure variable by itself - helps show which building materials had much or less severe damage

superstructure_summary <- data_cleaned %>%
  select(damage_grade, all_of(superstructure_vars)) %>%
  pivot_longer(
    cols = all_of(superstructure_vars),
    names_to = "variable",
    values_to = "present"
  ) %>%
  filter(present == 1) %>%
  count(variable, damage_grade) %>%
  group_by(variable) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

ggplot(superstructure_summary,
       aes(x = variable, y = prop, fill = damage_grade)) +
  geom_col(color = "white") +
  scale_fill_manual(
    values = c(
      "1" = "skyblue",
      "2" = "darkgoldenrod1",
      "3" = "red1"
    )
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
labs(
  title = "Damage Levels for Different Building Materials",
  x = "Building Material",
  y = "Percent of Buildings",
  fill = "Damage Grade"
) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Below we can see that each secondary use also contains different proportions for building damage type, with agriculture and police secondary uses having the most `damage_grade=3`

```{r plot secondary use}
# Plot detailed secondary use variables only - exclude has_secondary_use because it is the broad master flag

secondary_summary <- data_cleaned %>%
  select(damage_grade, all_of(setdiff(secondary_use_vars, "has_secondary_use"))) %>%
  pivot_longer(
    cols = -damage_grade,
    names_to = "variable",
    values_to = "present"
  ) %>%
  filter(present == 1) %>%
  count(variable, damage_grade) %>%
  group_by(variable) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

ggplot(secondary_summary,
       aes(x = variable, y = prop, fill = damage_grade)) +
  geom_col(color = "white") +
  scale_fill_manual(
    values = c(
      "1" = "skyblue",
      "2" = "darkgoldenrod1",
      "3" = "red1"
    )
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
labs(
  title = "Damage Grade by Secondary Use",
  x = "Secondary Use",
  y = "Percent of Buildings",
  fill = "Damage Grade"
) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Phase 3: Converting to a Binary Classification Problem

After performing EDA and considering different model approaches, we've decided to restructure this into a binary classification problem which classifies whether or not there would be damage rather than which of three damage grades to expect. We've grouped levels 2 and 3 of damage_grade together as this simplifies our classification problem in an appropriate manner given the ordinal nature of our classes. We determined it would be most important to predict simply whether there would be damage, instead of whether that damage was moderate to high, as understanding buildings which are at risk for damage is the key.

- **damaged:** Yes for damage_grade=1 (minimal)

- **damaged**: No for damage_grade=2,3 (moderate to high)

Below we see the new predictor and its distribution and still must strongly plan for class imbalance.

```{r add new damaged indicator and plot distribution}
data_cleaned <- data_cleaned |>
  mutate(damaged = factor(if_else(damage_grade %in% c(2, 3), "Yes", "No"), 
                          levels = c("Yes", "No")))

#remove old damage grade column
data_cleaned <- data_cleaned %>%
  select(-damage_grade)
```

```{r plot new damaged indicator distribution}
# Plot target distribution
ggplot(data_cleaned, aes(damaged, fill = damaged)) +
  geom_bar() +
  labs(
    title = "Distribution of Damaged",
    x = "Damaged",
    y = "Count"
  ) +
    scale_fill_manual(
            values = c(
              "No" = "skyblue",
              "Yes" = "red1"
            )
          )
```

# Phase 4: Test and Train Split

Our next steps are to split the data into an 80/20 test-train split using `damaged` for strata to ensure that each data set (test and train) have equal proportions of our damaged indicator.

```{r stratify split on damaged}
set.seed(420) 
data_split <- initial_split(data_cleaned, prop = 0.80, strata=damaged) 
train <- training(data_split) 
test <- testing(data_split)

#Setting up our resamples for cross-val using training set
resamples <- vfold_cv(train, v = 5, strata = damaged)
```

# Phase 5: Feature Selection

## Attempt #1: Lasso

As part of our attempt at using Lasso to reduce our feature set, we address the high cardinality of geo levels by grouping the rare categories of `geo_level_3_id` and `geo_level_3_id` which occur less than 1% into an "other" category. We take other preprocessing steps such as converting all categorical values to dummy numerics, removing any predictors that due not explain variance in the model, and normalizing all numeric predictors.

```{r lasso recipe prep}
# groups geographic levels together before dummy encoding
lasso_recipe <-
  recipe(damaged ~ ., data = train) %>%
#group rare categories <1% into "other"
#to reduce dummy variables/high cardinality
  step_other(
    geo_level_2_id,
    geo_level_3_id,
    threshold = 0.01
  ) %>%
#convert categorical to dummy numerics
  step_dummy(all_nominal_predictors()) %>%
#removes any variables without variance
#which indicates no model value due to no variance
  step_zv(all_predictors()) %>%
#scaled numerics predictors
  step_normalize(all_numeric_predictors())
```

Below we check the number of predictors after preprocessing to verify that dummy encoding & grouping of rare geographic levels did not create any large feature space.

```{r lasso predictor check}
prep(lasso_recipe) %>%
  juice() %>%
  dim()
```

The penalty controls how strongly LASSO shrinks coefficients toward 0. Larger penalties result in larger feature selection.

```{r lasso penalty grid}
# Create a grid of penalty values to check
lasso_grid <-
  grid_regular(
    penalty(),
    levels = 10
  )
```

We choose `mixture = 1` for L1 regularization, which can shrink coefficients to 0 and do automatic feature selection

```{r lasso specs}
# LASSO logistic reg model
library(glmnet)

lasso_spec <-
  logistic_reg(
    penalty = tune(), #seek best value by tuning
    mixture = 1
  ) %>%
  set_engine("glmnet")
```

Combine the preprocessing recipe/ model spec into a workflow.

```{r lasso wf}
lasso_wf <-
  workflow() %>%
  add_recipe(lasso_recipe) %>%
  add_model(lasso_spec)
```

Tune the penalty parameter using X val, evaluate model performance across penalty values and select the value that max ROC AUC.

```{r tune lasso penalty}
set.seed(420)

lasso_tuned <-
  tune_grid(
    lasso_wf,
    resamples = resamples, #5 CFV
    grid = lasso_grid, #10 penalty levels
    metrics = metric_set(roc_auc, bal_accuracy)
  )
```

With the below, we visualize how ROC AUC and balanced accuracy, which accounts for class imbalance, change as the penalty parameter get bigger.

```{r plot tuning of lasso}
autoplot(lasso_tuned)
```

Select the penalty value that produced the highest ROC AUC.

```{r get best penalty}
best_penalty <- select_best(
  lasso_tuned,
  metric = "roc_auc"
)

best_penalty
```

Fit the LASSO model using the best penalty value

```{r fit lasso with best penalty}
final_lasso_wf <-
  finalize_workflow(
    lasso_wf,
    best_penalty
  )

final_lasso_fit <-
  fit(
    final_lasso_wf,
    data = train
  )
```

Get coefficient estimates from the model at best penalty value

```{r}
coef_table <-
  tidy(
    extract_fit_parsnip(final_lasso_fit)
  )
```

Display the predictors ordered by coefficient magnitude which have not been filtered out by shrinking to 0.

```{r show predictors by coeff mag}
coef_table %>%
  filter(term != "(Intercept)") %>%
  arrange(desc(abs(estimate)))%>%
  filter(estimate != 0)
```

Display all features which have not had their coefficient shrunk to 0.

```{r get count of predictors with nonzero coeff}
coef_table %>%
  filter(estimate != 0)%>%
  nrow()
```

Per below, we have no actually had any features reduced to zero, merely specific values for those features.

```{r show which predictors are dropped}
coef_table %>%
filter(estimate == 0)
```

### LASSO Feature Selection Results

- ROC AUC: 0.895

- Best penalty: 1e-10

- Predictors retained: 89

- Conclusion:\
  LASSO provided little to no feature reduction implying most predictors appear to contain useful predictive information. Even when selecting a higher penalty value, such as 10\^-3, it resulted in minimal feature selection. We suspect this is because our data violates some of the assumptions of LASSO logistic regression; namely our sample size is much larger than our parameters, and many of our features are nested within one another and are not mutually exclusive. Because of this we will make a new attempt to determine feature importance using random forest models to perform feature selection.

## Attempt #2: Random Forest

For Random Forest, we chose permutation as our importance metric because it is model agnostic.

```{r}
library(tidymodels)
library(ranger)
library(vip)
library(important)

#Double check this is actually as factors and not dummies

set.seed(420)

#new df from the training data so need to mutate again
data_rf <- train %>%
  mutate(
    damaged = as.factor(damaged),
    across(
      c(
        geo_level_1_id,
        geo_level_2_id,
        geo_level_3_id,
        land_surface_condition,
        foundation_type,
        roof_type,
        ground_floor_type,
        other_floor_type,
        position,
        plan_configuration,
        legal_ownership_status
      ),
      as.factor
    )
  )

#recipe for random forest model
rf_recipe <-
  recipe(damaged ~ ., data = data_rf)

#pretty forest
rf_spec <-
  rand_forest(
    trees = 100
  ) %>%
  set_engine(
    "ranger",
    importance = "permutation"
  ) %>%
  set_mode("classification")

# workflow as usual
rf_wf <-
  workflow() %>%
  add_recipe(rf_recipe) %>%
  add_model(rf_spec)

# Fit the model

rf_fit <-
  rf_wf %>%
  fit(data = data_rf)

# Generate variable importance plot to determine what to keep
rf_fit %>%
  extract_fit_engine() %>%
  vip(num_features = 20)

```

Per our results, we see that `geo_level_1_id` dominates all other variables, followed by \*`mud_mortar_stone`, and other `geo_level_*_id`. We suspect the nested nature of geo levels is dominating the feature importance for random forest model, so choose to remove the finer granularity of `geo_level_2_id` and `geo_level_3_id` and keep ONLY `geo_level_1_id`.

Below we rerun random forest keeping ONLY `geo_level_1_id` in that training to assess the resulting feature importance.

```{r}
#code from various documents/tutorials
set.seed(420)
# Recipe with useless variables removed
rf_recipe <-
  recipe(damaged ~ ., data = train) %>%
  step_rm(
    geo_level_2_id,
    geo_level_3_id
  ) %>%
  step_zv(all_predictors()) #from one of the tutorials 

# Random Forest with permutation importance
rf_spec <-
  rand_forest(
    trees = 200
  ) %>%
  set_engine(
    "ranger",
    importance = "permutation",
    num.threads = parallel::detectCores() - 1 #speeds up
  ) %>%
  set_mode("classification")

# Workflow
rf_wf <-
  workflow() %>%
  add_recipe(rf_recipe) %>%
  add_model(rf_spec)

# Fit ONE model
final_rf_fit <-
  rf_wf %>%
  fit(data = train)

# Variable importance plot
final_rf_fit %>%
  extract_fit_engine() %>%
  vip(num_features = 20)

# numeric importance table for all variables
importance_tbl <-
  final_rf_fit %>%
  extract_fit_engine() %>%
  vip::vi()

importance_tbl %>%
  arrange(desc(Importance))

```

We now extract the features that account for 90% of the cumulative importance and are left with these 15 features.

```{r}
importance_tbl <- importance_tbl %>%
  arrange(desc(Importance)) %>%
  mutate(
    cum_importance = cumsum(Importance) / sum(Importance)
  )

selected_features <- importance_tbl %>%
  filter(cum_importance <= 0.90) %>%
  pull(Variable)

selected_features <- c(
  selected_features,
  importance_tbl %>%
    filter(cum_importance > 0.90) %>%
    slice(1) %>%
    pull(Variable)
)
selected_features <- unique(selected_features)

train_reduced <- train %>%
  select(any_of(selected_features), damaged)

test_reduced <- test %>%
  select(any_of(selected_features), damaged)

```

### Random Forest Feature Selection Results

After we removed hierarchical variables `geo_level_2_id` and `geo_level_3_id`, we selected variables that were most important and accounted for 90% of total importance in the data. This preserves most of the future model's predictive performance while simplifying model interpretation and workflow. Random Forest has yielded a more productive result for us than Lasso.

# Phase \# 5 Helper Function Definitions

We establish the set of metrics to evaluate each model against to maintain consistency and use helper functions to ensure accurate calculation against each model with minimal code redundancy.

```{r define metric set and helper functions}
#get_model_metrics(your fitted model before augment against test data,"Your model name")
#usage example: 
#get_model_metrics(final_lasso_fit,"Lasso")
#this will yield both the metrics table AND the ROC AUC Plot

#define metric set to be used across models for evaluations and comparison
class_metrics<- metric_set(roc_auc, bal_accuracy, sens, spec, f_meas, kap)

#pass fitted model using best parameters + model name
#function will predict against test data
get_model_preds <- function(fitted_model, model_name, new_data) {
  augment(fitted_model, new_data = new_data) ## !!!IMPORTANT" <- if we are going to train against train_reduced than this line needs to be changed to new_data=test_reduced. right now it say test because both RF and Lasso were trained against train in full
}

#separate visuals out for organization
get_model_plots <- function(aug_data, model_name) {
  roc_plot<-aug_data |>
    roc_curve(truth = damaged, .pred_Yes) |>
    autoplot() +
    labs(
      title = paste0(model_name, " ROC Curve"),
      x = "False Positive Rate",
      y = "True Positive Rate"
    )
}


get_model_metrics <- function(fitted_model, model_name, new_data) {

  #get predicted model data to send to graphs
  aug_data<-get_model_preds(fitted_model, model_name, new_data)

  #use same table for metrics table
  metrics_results<-aug_data|>
    class_metrics(truth = damaged, estimate = .pred_class, .pred_Yes) |>
    mutate(model_name = model_name) |>
    select(model_name, .metric, .estimate) |>
    pivot_wider(names_from = .metric, values_from = .estimate)
   
   #show graphics
   print(get_model_plots(aug_data, model_name)) 
   return(metrics_results)
}
```

# SAMPLE HELPER FUNCTION USAGE ONLY!! TO BE REMOVED AND ONLY USED FOR MODELS.

```{r lasso evaluation}
get_model_metrics(final_rf_fit,"Random forest Feature Selection Only", test)|>
  knitr::kable(digits = 4)
get_model_metrics(final_lasso_fit,"Lasso Feature Selection Only", test)|>
  knitr::kable(digits = 4)
```

# Models

Now that we have a reduced feature list from our random forest model, we can now attempt to build models off of our new reduced data set, train_reduced. We break out models into 5 major sections; Logistic Regression, LDA, QDA, Random Forests with Ensemble Methods, and XGBoost. 

## Logistic Regression with L1 (LASSO) and L2 (Ridge) Regularization

We will start this analysis in threes, one with pure lasso (after feature selection), one with pure ridge, and one with a mix of both, which is an elasticnet model. 

### Pure L1

```{r create tunning grid}
logistic_grid <-
  grid_regular(
    penalty(),
    levels = 10
  )
```

```{r Log Reg}

lasso_spec <-
  logistic_reg(
    penalty = tune(),
    mixture = 1        # L1 Regularization
  ) %>%
  set_engine("glmnet")

#New recipe based on the results from feature selection
logistic_recipe <-
  recipe(damaged ~ ., data = train_reduced) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors())

#New CV folds for reduced dataset
resamples_reduced <- vfold_cv(train_reduced, v = 5, strata = damaged)

lasso_wf <-
  workflow() %>%
  add_recipe(logistic_recipe) %>%
  add_model(lasso_spec)

set.seed(42)

lasso_tuned <-
  tune_grid(
    lasso_wf,
    resamples = resamples_reduced,
    grid = logistic_grid,
    metrics = metric_set(roc_auc, bal_accuracy)
  )
```

```{r View tuning results}
autoplot(lasso_tuned)
```

```{r select best penalty}
best_lasso <-
  select_best(
    lasso_tuned,
    metric = "roc_auc"
  )
```
```{r Fin WF}
final_lasso_wf <-
  finalize_workflow(
    lasso_wf,
    best_lasso
  )
```
```{r FM}
final_lasso_fit <-
  fit(
    final_lasso_wf,
    data = train_reduced
  )
```
```{r Review LASSO coefficents}
lasso_coef <-
  tidy(
    extract_fit_parsnip(final_lasso_fit)
  )

lasso_coef %>%
  filter(term != "(Intercept)") %>%
  arrange(desc(abs(estimate)))
```

### Pure L2

```{r Log Re Ridgeg}
ridge_spec <-
  logistic_reg(
    penalty = tune(),
    mixture = 0        # L2 Regularization
  ) %>%
  set_engine("glmnet")

ridge_wf <-
  workflow() %>%
  add_recipe(logistic_recipe) %>%
  add_model(ridge_spec)

set.seed(42)

ridge_tuned <-
  tune_grid(
    ridge_wf,
    resamples = resamples_reduced,
    grid = logistic_grid,
    metrics = metric_set(roc_auc, bal_accuracy)
  )
```
```{r Logistic Reg Ridge continued}
# View tune results
autoplot(ridge_tuned)

# Select best penalty
best_ridge <-
  select_best(
    ridge_tuned,
    metric = "roc_auc"
  )

# Finalize wf
final_ridge_wf <-
  finalize_workflow(
    ridge_wf,
    best_ridge
  )

# Fit model
final_ridge_fit <-
  fit(
    final_ridge_wf,
    data = train_reduced
  )
```
```{r Log Reg Ridge coefficents}
ridge_coef <-
  tidy(
    extract_fit_parsnip(final_ridge_fit)
  )

ridge_coef %>%
  filter(term != "(Intercept)") %>%
  arrange(desc(abs(estimate)))
```

### L1 and L2 (Elastic)

Sources for running this code in a semi-speedy manner can be found in this documentation: https://tune.tidymodels.org/reference/control_bayes.html and https://tune.tidymodels.org/reference/tune_bayes.html

```{r elastic log reg}
set.seed(420)

# specify the elastic model:
elastic_spec <-
  logistic_reg(
    penalty = tune(),
    mixture = tune()
  ) %>%
  set_engine("glmnet") %>%
  set_mode("classification")

#set up my param space
elastic_params <-
  parameters(
    penalty(),
    mixture())

elastic_params <- finalize(elastic_params, train_reduced)

#wf like normal
elastic_wf <-
  workflow() %>%
  add_recipe(logistic_recipe) %>%
  add_model(elastic_spec)

#bayesian hyperparameter optimization
#control aspects found in tidyverse documentation
elastic_tune <-
  tune_bayes(
    elastic_wf,
    resamples = resamples_reduced,
    initial = 10,
    iter = 15,
    param_info = elastic_params,
    metrics = metric_set(roc_auc,bal_accuracy),
    control = control_bayes(verbose = TRUE, no_improve = 5, save_pred = FALSE)
  )

#Grab our best version
best_elastic <-
  select_best(elastic_tune, metric = "roc_auc")

best_elastic

```

According to this, our best elasticnet parameters uses a penalty of 0.0000359 and a mixture of 0.111.

```{r fit elastic log reg}
# Finalize wf
final_elastic_wf <-
  finalize_workflow(
    elastic_wf,
    best_elastic
  )

# Fit model
final_elastic_fit <-
  fit(
    final_elastic_wf,
    data = train_reduced
  )
```
```{r Log Reg elastic coefficents}
elastic_coef <-
  tidy(
    extract_fit_parsnip(final_elastic_fit)
  )

elastic_coef %>%
  filter(term != "(Intercept)") %>%
  arrange(desc(abs(estimate)))
```


### Summary

```{r test metrics}
get_model_metrics(
  final_lasso_fit,
  "L1 Log Regression -LASSO",
  test_reduced
) |>
  knitr::kable(digits = 4)

get_model_metrics(
  final_ridge_fit,
  "L2 Log Reg - Ridge",
  test_reduced
) |>
  knitr::kable(digits = 4)

get_model_metrics(
  final_elastic_fit,
  "L1+L2 Log Reg - Elastic",
  test_reduced
) |>
  knitr::kable(digits = 4)
```

Here we see that the model metrics show that, in terms of test metrics, all three of these models are roughly balanced. Elastic net based model does the best overall, with the highest ROC_AUC and the best balanced accuracy. That being said, these are the general metrics looking at *all* responses. We care about predicting whether or not a building will get damage, and damage is a rare class. Therefore, we should also consider the f_meas and sensitivity. In terms of sensitivity scores, the ridge regression does best, having the sensitvity of 0.9803. the other two models are tied at 0.9778, which is not that far off. In terms of f_meas, the elastic net and lasso are tied for the best at an f_meas of 0.9533. 

Over all, these three differently trained models are very similar in terms of test performance, with the elastic model slightly winning out overall. 

Let's check training metrics for any signs of overfitting:


```{r train metrics}
get_model_metrics(
  final_lasso_fit,
  "L1 Log Regression -LASSO",
  train_reduced
) |>
  knitr::kable(digits = 4)

get_model_metrics(
  final_ridge_fit,
  "L2 Log Reg - Ridge",
  train_reduced
) |>
  knitr::kable(digits = 4)

get_model_metrics(
  final_elastic_fit,
  "L1+L2 Log Reg - Elastic",
  train_reduced
) |>
  knitr::kable(digits = 4)
```

No real signs of overfitting, that's good!